---
title: "Data Calculations"
author: "Jason M. Gleditsch^[Temple University, jmg5214@gmail.com]"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: TRUE
    toc_depth: 3
  pdf_document:
    toc: TRUE
    toc_depth: 3
---

_**HI I CHANGED SOMETHING**_

# Aim and Scope

Here we calculate the necessary data for the Island Biogeographic Analyses of the Caribbean Herpetofauna. We will be determining the species richness (SR) and phylogenetic species variability (PSV) for each island bank in the Caribbean. Each of these values are going to be calculated for various species groups (i.e. clades) as determined by the taxonomic hierarchies for the reptile and amphibian species present in the Caribbean Bioregion as delineated by Hedges et al. (2019) and including Bermuda.

The SR will also be calculated for clades that are delineated by using the phylogeny so that the group of species that are descendant from each node are a clade.

First we need to load in the `caribmacro` package.

```{r project pkg, message = FALSE}
# Project package
library(caribmacro)

```

If you do not have this package installed, then run `devtools::load_all` in order to load all of the necessary functions if you are running this code in the caribmacro R project in the shared folder. If you are not, then see the functions vignette for the Functions needed.

```{r project functions, eval = FALSE, message = FALSE}
# Load all of the functions of the caribmacro package
devtools::load_all("./", export_all = FALSE)

## Function needed ##
#   'hierarchy'  -> returns the taxonomic hierarcy
#   'com_matrix'  -> determines the communities within geographic features
#   'SR_geo'  -> calculates the species richness within geographic features
#   'PSV_geo'  -> calculates the psv within geographic features
#   'clade_age'  -> calculates the crown age of clades
#   'node_clades'  -> determines the clades of species based on phylogenetic nodes
#   'clade_area'  -> determines the number of geographic feature each clade is found on
#   'row_max'  -> determines the maximum value in each row of a data frame
#   'factorize'  -> turns character columns of a data frame into factors
```

The other R packages needed for this are:

```{r required packages, message = FALSE}
library(here)
library(reshape)
library(readr)
library(pwr)

# Spatial Packages
library(sf)
library(rgeos)

# Analytical Packages
library(FactoMineR)

# Diversity Database Packages
library(taxize)

# Phylogenetic Packages
library(ape)
library(phyr)

# Graphing Packages
library(ggplot2)
library(gridExtra)
```

# Island Banks

For our analyses we will be clustering the islands in the Caribbean into banks based on Powell and Henderson 2012 and Hedges et al. 2019. To further delineate banks we used the General Bathymetric Chart of the Oceans (GEBCO Compilation Group 2020, downloaded September 2020) to group islands based on the under water topography.

```{r shp load, include = FALSE}
mainland <- st_read(file.path(here(), 'data_raw', 'gis', 'Cont_Main_WGS_Merc.shp'))
carib.bnk <- st_read(file.path(here(), 'data_raw', 'gis', 'Carib_Banks.shp'))
```

```{r shp projection, include = FALSE}
# Re-project into Lambert Azimuthal Equal Area centered in the Caribbean
mainland <- st_transform(mainland, "+proj=laea +x_0=0 +y_0=0 +lon_0=-72 +lat_0=20")
carib.bnk <- st_transform(carib.bnk, "+proj=laea +x_0=0 +y_0=0 +lon_0=-72 +lat_0=20")

# Change 'mainland' and 'carib.bnk' to class 'Spatial'
mainland <- as(mainland, Class = "Spatial")
carib.bnk <- as(carib.bnk, Class = "Spatial")

```

```{r bank mcp, include = FALSE}
bnk.mcp <- gConvexHull(carib.bnk, byid = TRUE)

bnk.mcp.sf <- st_as_sf(bnk.mcp)
bnk.mcp.sf$bank <- st_as_sf(carib.bnk)$bank
```

Let's now take a look at the banks.

```{r mcp plot, echo = FALSE, fig.height = 7, fig.width = 7, fig.align="center"}
ggplot() +
    geom_sf(data = bnk.mcp.sf, fill = 'red', color = 'black') +
    geom_sf(data = st_as_sf(carib.bnk), fill = 'azure2', color = 'grey20') +
    geom_sf(data = st_as_sf(mainland), fill = 'azure2', color = 'grey20') +
    xlab('Longitude') + 
    ylab('Latitude') +
    labs(title = 'Banks with their AOE', 
         subtitle = expression('(Red = Area of extent; Gray = Bank)'))

```


# Taxonomic Clade Communities

## Community Attributes

Now that we have determined bank area, we can now start working on the Community attributes. These attributes are their species richness (SR) and phylogenetic species variability (PSV). 

### Species Hierarchy

We will calculate SR and PSV for various clades of species on each bank. Therefore, we need to determine the species hierarchies first. We will do this using a function in the 'caribherp' package. This function is `hierarchy()` and uses the package 'taxize' to look up and download species hierarchies. The function in 'taxize' it uses makes mistakes, and therefore, we need to check any errors outside of R. Once any errors are fixed we then load the data back in. When we checked our species hierarchies we saw that a couple species names were matched with an insect or a plant species. The rest of the errors were all due to misspellings or ids that were only to the genus level.

For this we need to read in the species occurrence data in the 'data_raw' folder.

```{r records load}
herp <- read.csv(file.path(here::here(), 'data_raw', 'IBT_Herp_Records_final_SR_Drivers.csv'), header=TRUE)
```

With the species occurrence data we can now run the `hierarchy()` function and save the resulting data.
```{r hierarchies, eval = FALSE}
carib.hier <- hierarchy(data = herp, species = 'binomial', db = 'gbif')

#Need to check/correct any errors and missing hierarchies in Excel
write.csv(carib.hier, file.path(here(), 'data_raw', 'supporting_data', 'IBT_Carib_Sp_Taxonomy_v0.csv'), row.names=FALSE)
```

Remember, we need to check the data outside R for errors and fix those errors. Once we do that we can then load in the new fixed hierarchies and merge it with the occurrence data.

```{r hierarchies read}
carib.hier <- read.csv(file.path(here(), 'data_raw', 'supporting_data', 'IBT_Carib_Sp_Taxonomy_v1.csv'), header=TRUE)

# Combine Hierarchies with Occurrence Data
names(carib.hier)[which(names(carib.hier) == 'sp_in_Data')] <- 'binomial'

herp <- merge(herp, carib.hier, by = 'binomial', all = TRUE)

```

Let's check once more to see if we are missing any hierarchies

```{r hierarchy check}
# Check for species missing their hierarchy
herp$binomial <- as.factor(herp$binomial)
chk <- levels(droplevels(herp[is.na(herp$Genus), 'binomial']))
chk
```

We can see that there are no missing hierarchies in our species occurrence data. However, the `hierarchy()` function only finds a species "typical" taxonomic hierarchy. If we want to add in certain groups that are not a kingdom, phylum, class, order, family, or genus, than we need to add them manually. Because we are interested in all snakes, geckos, and skinks , we will include the species' suborders or smaller groups so that we include groups of interest. For instance, the non-Iguania squamates were broken up into the infraorders Gekkota, Amphisbaenia, Scincomorpha, and Anguimorpha (which combines the infraorders Diploglossa and Platynota in our data) and the superfamily Gymnophthalmoidea. 

The suborder/infraorder groups were determined by searching the species' families on the Integrated Taxonomic Information System (ITIS) and then doing a web search if the family's ITIS standard report page did not list a suborder/infraorder.

However, the searches on ITIS illuminated a potential error in the identification of the Natricidae family. The family has is typically been considered a subfamily of Colubridae (Pryon et al. 2011). Therefore, we will change this family to match ITIS.

```{r more groups}
# Fix Family Error
herp[which(herp$Family=='Natricidae'), 'Family'] <- 'Colubridae'

# Add suborder groups
sorder <- read.csv(file.path(here(), 'data_raw', 'supporting_data', 'IBT_Carib_Suborders.csv'), header=TRUE)

herp <- merge(herp, sorder, by = 'Family', all.x = TRUE)
carib.hier <- merge(carib.hier, sorder, by = 'Family', all.x = TRUE)

chk <- levels(droplevels(as.factor(herp[is.na(herp$Suborder), 'Family'])))
chk
```


### Community Matrices

Now that we have the hierarches for each species in our occurrence data, we can create the community matrices for each clade. These matrices will comprise of j columns, where j equals the number of species in the clade, and k rows, where k equals the number of banks the clade is recorded on. Each cell is then populated by a 1 or 0, where 1 represents the presence of the jth species on the kth bank, and 0 represents its absence.

In order to create these matrices we will use the `com_matrix()` function from the 'caribmacro' package. This function returns a named list of i community matrices where i is the number of species groups. In our case we will create a community matrix for all herpetofuana and all of the classes, orders, families, and genuses in our occurrence data. Additionally, we will break up each of these clades based on thier status on the banks (i.e. native or exotic) and the entire clade. This comes out to be 693 species groups.

```{r community matrices}
# First simplify bank status
herp$stat_new <- NA
for (i in 1:nrow(herp)) {
  if (!is.na(herp[i, 'bnk_status'])) {
    if (herp[i, 'bnk_status'] == 'E') {
      herp[i, 'stat_new'] <- 'E'
    } else if (herp[i, 'bnk_status'] == 'FE') {
      herp[i, 'stat_new'] <- 'N'
    } else if (herp[i, 'bnk_status'] == 'N') {
      herp[i, 'stat_new'] <- 'N'
    } else if (herp[i, 'bnk_status'] == 'PX') {
      herp[i, 'stat_new'] <- 'N'
    } else if (herp[i, 'bnk_status'] == 'U') {
      herp[i, 'stat_new'] <- 'E'
    } else {
      herp[i, 'stat_new'] <- 'X'
    }
  } else {
    herp[i, 'stat_new'] <- 'X'
  }
}

# Fix specific statuses
## Considered extinct
herp[which(herp$binomial == 'Crocodylus rhombifer' & herp$bank == 'grand cayman'), 'stat_new'] <- 'X' 
## Considered introduced by caribherp.org
herp[which(herp$binomial == 'Chelonoidis carbonarius' & herp$bank == 'anguilla'), 'stat_new'] <- 'E' 

           


system.time(coms <- com_matrix(data = herp,
                   species = "binomial",
                   geo_group = "bank",
                   taxa_group = c("Genus", "Family", "Suborder", "Order", "Class"),
                   status = "stat_new",
                   stat_levels = c("N", "E"),
                   total = TRUE))

# List the statuses excluded from the analyses
levels(as.factor(herp[which(herp$stat_new == 'X'), 'bnk_status']))

```

For our analyses we are changing 'unknown' (U) statuses to 'exotic' (E) and the 'possibly extinct' (PE) and 'have fossil but extant' (FE) statuses to 'native' (N). We are also excluding 'only known by fossils' (F), 'waifs' (W), 'extirpated exotics' (EI), and records only identified to the genus level but the genus is found elsewhere on the bank (G).

Because our the way we are grouping species hierarchical, we may have clades that have the same species (e.g. the suborder clade Eusuchia has all of the same species as the order Crocodilia). Let's check to see if this occurs in our data by looking at how many clades there are in the each level downstream from a clade. Those clades that only have one clade in a lower level will have the same species as the lower level.

```{r clade rem}
tst <- herp
tst$binomial <- as.factor(tst$binomial)

grps <- c("Class", "Order", "Suborder", "Family", "Genus")

lists <- vector('list', length(grps))
names(lists) <- grps

for (i in 1:length(grps)) {
  tst[, grps[i]] <- as.factor(tst[, grps[i]])
  lists[[grps[i]]] <- levels(tst[, grps[i]])
}

out <- data.frame(NULL)
for (i in 1:length(lists)) {
  if (i != length(lists)) {
    for (j in 1:length(lists[[i]])) {
      temp <- droplevels(tst[which(tst[, names(lists)[i]] == lists[[i]][j]), 
                             c('binomial', grps[(i+1):length(grps)])])
      
      for (k in 2:ncol(temp)) {
        if (length(levels(temp[, k])) > 1) {
          lvls <- levels(temp[, k])
          tmp <- data.frame(Group.1 = rep(lists[[i]][j], length(lvls)),
                            Grp.1.lvl = rep(grps[i], length(lvls)),
                            Group.2 = lvls,
                            Grp.2.lvl = rep(which(grps == names(temp)[k]), length(lvls)),
                            Different = rep(TRUE, length(lvls)))
          
          out <- rbind(out, tmp)
        } else {
          lvls <- levels(temp[, k])
          tmp <- data.frame(Group.1 = rep(lists[[i]][j], length(lvls)),
                            Grp.1.lvl = rep(i, length(lvls)),
                            Group.2 = lvls,
                            Grp.2.lvl = rep(which(grps == names(temp)[k]), length(lvls)),
                            Different = rep(FALSE, length(lvls)))
          
          out <- rbind(out, tmp)
        }
      }
    }
  } else {NA}
}
rm(tst, temp, tmp, lists, grps, lvls)

# To check that the code work:
# View(out[!out$Different, ])

rem <- levels(as.factor(out[!out$Different, 'Group.1']))
cat(paste0('Clades removed (n = ', length(rem), ') \n', '\n'), paste(rem, collapse = " | "))

```

As you can see there are a lot of clades that contain the same species as younger clades. Because this would be psuedoreplication (i.e. the same species group analyzed twice), we need to remove these clades from our analyses.

```{r com rem}
good <- gsub('.T', '', names(coms)[grep('.T', names(coms), fixed = TRUE)], fixed = TRUE)

good <- paste(good, ".", sep = "")
rem <- paste(rem, ".", sep = "")

good <- setdiff(good, rem)

tmp <- coms[names(coms)[grep(good[1], names(coms), fixed = TRUE)]]
for (i in 2:length(good)) {
  tmp <- c(tmp, coms[names(coms)[grep(good[i], names(coms), fixed = TRUE)]])
}

coms <- tmp
rm(good, tmp, rem)
```


### Species Richness Calculation

Now that we have the community matrices we can calculate the species richness (SR) of each species group on each bank. To do this we will use the `SR_geo()` function in the 'caribmacro' package.

```{r species richness}
system.time(SR <- SR_geo(data = coms, geo_group = "bank"))
SR_out <- SR
```

### Clade Species Richness

We may also want to include the species richness of each clade in our analysis, because the clades may differ in their speciation rates. We will only calculate this for the Caribbean Bioregion. Therefore, the SR of each clade will represent the rate a clade will add a species due to speciation (for the native species richness) or both speciation and species introduction (both native and exotic species richness) in the Caribbean Bioregion.

We will calculate clade SR in a similar way to bank SR (see above). However, instead of setting `geo_group = "bank"`, we will set `geo_group = "region"` in the `com_matrix()` function, where 'region' is set to 'carib' representing the Caribbean.

```{r clade SR}
herp$region <- 'carib'
clade.coms <- com_matrix(data = herp,
                         species = "binomial",
                         geo_group = "region",
                         taxa_group = c("Genus", "Family", "Suborder", "Order", "Class"),
                         status = "stat_new",
                         stat_levels = c("N", "E"),
                         total = TRUE)

clade.SR <- SR_geo(data = clade.coms, geo_group = "region")
tmp <- strsplit(names(clade.SR)[which(names(clade.SR) != 'region')], ".", fixed = TRUE)
tmp <- data.frame(matrix(unlist(tmp), nrow = length(tmp), byrow = TRUE))

clade.SR <- cbind(tmp, as.numeric(clade.SR[1, names(clade.SR)[which(names(clade.SR) != 'region')]]))
names(clade.SR) <- c('Taxon', 'Group', 'SR')

```


### Clade P/A Matrix

It may be important to have a matrix that has the presence and absence of species in each clade

```{r p_a matrix}
sp <- levels(as.factor(herp$binomial))
sp <- gsub(' ', '_', sp, fixed = TRUE)

cld.mat <- as.data.frame(matrix(NA, nrow(clade.SR), length(sp)))
names(cld.mat) <- sp
cld.mat$Clade <- clade.SR$Taxon
cld.mat$Group <- clade.SR$Group

cld.mat <- cld.mat[, c('Clade', 'Group', sort(names(cld.mat)[which(names(cld.mat) != 'Clade' & 
                                                                     names(cld.mat) != 'Group')]))]

for (i in 1:nrow(cld.mat)) {
  if (cld.mat[i, 'Group'] != 'T') {
    if (is.element(cld.mat[i, 'Clade'], herp$Class)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Class == cld.mat[i, 'Clade'] & herp$stat_new == cld.mat[i, 'Group']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
    
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Order)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Order == cld.mat[i, 'Clade'] & herp$stat_new == cld.mat[i, 'Group']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff( names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Suborder)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Suborder == cld.mat[i, 'Clade'] & herp$stat_new == cld.mat[i, 'Group']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Family)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Family == cld.mat[i, 'Clade'] & herp$stat_new == cld.mat[i, 'Group']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Genus)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Genus == cld.mat[i, 'Clade'] & herp$stat_new == cld.mat[i, 'Group']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
    
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else {NA}
  } else {
    if (is.element(cld.mat[i, 'Clade'], herp$Class)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Class == cld.mat[i, 'Clade']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Order)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Order == cld.mat[i, 'Clade']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Suborder)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Suborder == cld.mat[i, 'Clade']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Family)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Family == cld.mat[i, 'Clade']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else if (is.element(cld.mat[i, 'Clade'], herp$Genus)) {
      tmp.sp <- levels(as.factor(droplevels(herp[which(herp$Genus == cld.mat[i, 'Clade']), 'binomial'])))
      tmp.sp <- gsub(' ', '_', tmp.sp, fixed = TRUE)
      
      cld.mat[i, tmp.sp] <- 1
      cld.mat[i, setdiff(names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')], tmp.sp)] <- 0
    } else {NA}
  }
}

cld.mat[which(cld.mat$Clade == 'All'), names(cld.mat)[which(names(cld.mat) != 'Clade' & names(cld.mat) != 'Group')]] <- 1
```

## Data Saving

Now lets save all of the community and clade attributes we calculated in the 'data_out' folder. We also want to save the clade SR and the Caribbean species taxonomy in the supp_info folder.

```{r data saving}
# Data for data_out
write.csv(SR_out, file.path(here(), 'data_out', 'sr_drivers', 'SR_Drivers_Herp_SR.csv'), row.names = FALSE)
write.csv(clade.SR, file.path(here(), 'data_out', 'sr_drivers', 'SR_Drivers_Clade_SR.csv'), row.names = FALSE)

# Data for supp_info
write.csv(carib.hier, file.path(here(), 'data_out', 'sr_drivers', 'supp_info', 'SR_Drivers_Carib_Taxonomy.csv'), row.names=FALSE)
write.csv(cld.mat, file.path(here(), 'data_out', 'sr_drivers', 'supp_info', 'SR_Drivers_Taxonomy_Clade_Matrix.csv'), row.names=FALSE)
```

```{r supp info save, include = FALSE}
clds <- c("All", "Reptilia", "Squamata", "Iguania", "Serpentes", "Anolis", "Gekkota", "Gymnophthalmoidea", "Typhlopidae", 
          "Testudines", "Iguanidae", "Cryptodira", "Amphibia", "Neobatrachia", "Eleutherodactylus", "Hylidae")

out <- NULL
for (i in 1:length(clds)) {
  tmp <- cbind(SR_out$bank, SR_out[, grep(clds[i], names(SR_out), fixed = TRUE)])
  names(tmp) <- c('Bank', 'Introduced_SR', 'Native_SR', 'Total_SR')
  if (clds[i] == 'All') {
    tmp$Clade <- 'All Herpetofauna'
  } else {
    tmp$Clade <- clds[i]
  }
  tmp <- tmp[, c('Clade', 'Bank', 'Introduced_SR', 'Native_SR', 'Total_SR')]
  out <- rbind(out, tmp)
}

write.csv(out, file.path(here(), 'data_out', 'results', 'sr_drivers', 'supp_info', 'Gleditsch_et_al_Caribbean_Herp_SR.csv'), row.names = FALSE)
```
